\documentclass[mistar]{acmtrans2m}
\usepackage{cite}
\usepackage{graphicx}

% Os seguintes packages são necessários caso escreva em português
\usepackage[latin1]{inputenc}
\usepackage[T1]{fontenc}
%

\newtheorem{theorem}{Theorem}[section]
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newdef{definition}[theorem]{Definition}
\newdef{remark}[theorem]{Remark}

%Altere os seguintes dois parâmetros.
%Os parâmetros, normalmente nome do autor no primeiro e  título do artigo no segundo, serão inseridos nos cabeçalhos das páginas pares e impares respectivamente.
\markboth{Nome}{Artigos para MI-STAR}

%Nome do Artigo
\title{Analysing and Mesuring Open-Source projects.}
 
 %Autor, em maiúsculas, e instituição, neste caso Universidade do Minho
\author{MIGUEL REGEDOR\\Universidade do Minho}
            
\begin{abstract} 
Nowadays, open-source software is well disseminated, thousands of software packages can be found in platforms like Github or Sourceforge.

Those platforms (usually associated with CVS repositories) make it easy for developers, all around the globe, to collaborate in open-source projects, and also work as a way to make software available to users. But, when users are looking for an open source project, they have no information about the quality of the project: does it have good documentation? does the code follow standards? is it similar to other projects? Few users have the ability to answer this questions by looking at the code.
We believe that an application able to do an automatic analysis of a software package and to generate a hight level overview would enable users to make better choices. It would also help developers by giving hints on how to further improve their software.

In this context, this paper discuss what should be taken into account to develop such an application with the focus on Github projects, and what benefits can be expected for Github users and developers.
\end{abstract}
 
%Categorize o artigo conforme as categorias da ACM.
%Alguns exemplos de categorias. 
\category{D.2.8}{Software Engineering}{Metrics}
\category{F.3.0}{Logics and meanings of programs}{General}
\category{I.5.0}{Pattern Recognition}{General}

%Insira um, ou mais, dos seguintes termos em \terms
%Algorithms; Design; Documentation; Economics; Experimentation; Human factors; Languages; Legal aspects; Management; Measurement; Performance; Reliability; Security; Standardization; Theory; Veri?cation; 
\terms{Documentation, Standardization} 
  
    
\keywords{Static code analysis, Open-Source, Artigos, MI-STAR, Universidade do Minho}
            
\begin{document}
\setcounter{page}{1}%Alterar "1" para a página a ser indicada pela equipa do MI-STAR.
            
\begin{bottomstuff} 
Inserir aqui informações adicionais acerca do artigo e autor. Tipicamente contactos do autor. 
\end{bottomstuff}
            
\maketitle

\section{Introdução}
...
\section{Uma secção}
....
\subsection{Subsecção}
...
\subsection{Outra subsecção}
...
\section{Conclusão}
...

%Bibliografia
\bibliography{bibtext}%substituir o parâmetro "bibtext" pelo nome do ficheiro (sem extensão) que contém a bibliografia BibTex
\bibliographystyle{acmtrans}


\begin{received}
\end{received}
\end{document}


