\thispagestyle{empty}
\appendix
\chapter{Rails Best Practices}\label{app:rails_best_practices}


\section{Remove Tab}
  Make sure there are no tabs in files.
 
  % See the best practice details here \url{http://rails-bestpractices.com/posts/81-remove-tab}
   
\section{Remove Trailing Whitespace }  
  Make sure there are no trailing whitespace in codes.
 
  % See the best practice details here \url{http://rails-bestpractices.com/posts/60-remove-trailing-whitespace}
   
\section{Add Model Virtual Attribute}
  Make sure to add a model virual attribute to simplify model creation.
 
  % See the best practice details here \url{http://rails-bestpractices.com/posts/4-add-model-virtual-attribute}
       
\section{Always Add Db Index}
  Review db/schema.rb file to make sure every reference key has a database index.
 
  % See the best practice details here \url{http://rails-bestpractices.com/posts/21-always-add-db-index}
       
\section{Dry Bundler In Capistrano}
  Review config/deploy.rb file to make sure using the bundler's capistrano recipe.
 
  % See the best practice details here \url{http://rails-bestpractices.com/posts/51-dry-bundler-in-capistrano}
   
\section{Isolate Seed Data }
  Make sure not to insert data in migration, move them to seed file.
 
  % See the best practice details here \url{http://rails-bestpractices.com/posts/20-isolating-seed-data.}
   
\section{Keep Finders On Their Own Model}
  Review model files to ake sure finders are on their own model.
 
  % See the best practice details here \url{http://rails-bestpractices.com/posts/13-keep-finders-on-their-own-model.}
   
\section{Law Of Demeter }
  Review to make sure not to avoid the law of demeter.
 
  % See the best practice details here \url{http://rails-bestpractices.com/posts/15-the-law-of-demeter.}
   
\section{Move Code Into Controller}
  Review a view file to make sure there is no finder, finder should be moved to controller.
 
  % See the best practice details here \url{http://rails-bestpractices.com/posts/24-move-code-into-controller.}
   
\section{Move Code Into Helper }
  Review a view file to make sure there is no complex options\_for\_select message call.
 
  % See the best practice details here \url{http://rails-bestpractices.com/posts/26-move-code-into-helper.}
   
\section{Move Code Into Model }
  Review a view file to make sure there is no complex logic call for model.
 
  % See the best practice details here \url{http://rails-bestpractices.com/posts/25-move-code-into-model.}
     
\section{Move Finder To Named Scope }
  Review a controller file to make sure there are no complex finder.
 
  % See the best practice details here \url{http://rails-bestpractices.com/posts/1-move-finder-to-named_scope.}
   
\section{Move Model Logic Into Model}
  Review a controller file to make sure that complex model logic should not exist in controller, should be moved into a model.
 
  % See the best practice details here \url{http://rails-bestpractices.com/posts/7-move-model-logic-into-the-model.}
   
\section{Needless Deep Nesting}
  Review config/routes.rb file to make sure not to use too deep nesting routes.
 
  % See the best practice details here \url{http://rails-bestpractices.com/posts/11-needless-deep-nesting.}
   
\section{Not Use Default Route}
  Review config/routes file to make sure not use default route that rails generated.
 
  % See the best practice details here \url{http://rails-bestpractices.com/posts/12-not-use-default-route-if-you-use-restful-design}
   
\section{Not Use Time Ago In Words}
  Review view and helper files to make sure not use time\_ago\_in\_words or distance\_of\_time\_in\_words\_to\_now.
 
  % See the best practice details here \url{http://rails-bestpractices.com/posts/105-not-use-time_ago_in_words.}
   
\section{Overuse Route Customizations}
  Review config/routes.rb file to make sure there are no overuse route customizations.
 
  % See the best practice details here \url{http://rails-bestpractices.com/posts/10-overuse-route-customizations.}
   
\section{Protect Mass Assignment}
  Review model files to make sure to use attr\_accessible or attr\_protected to protect mass assignment.
 
  See the best practices details here \url{http://rails-bestpractices.com/posts/148-protect-mass-assignment.}
   
\section{Remove Empty Helpers Review}
  Review a helper file to make sure it is not an empty moduel.
 
  % See the best practice details here \url{http://rails-bestpractices.com/posts/72-remove-empty-helpers.}
   
\section{Remove Unused Methods In Controllers}
  Find out unused methods in controllers.
   
\section{Remove Unused Methods In Helpers}
  Find out unused methods in helpers.
       
\section{Remove Unused Methods In Models}
  Find out unused methods in models.
   
\section{Replace Complex Creation With Factory Method}
  Review a controller file to make sure that complex model creation should not exist in controller, should be replaced with factory method.
 
  % See the best practice details here \url{http://rails-bestpractices.com/posts/6-replace-complex-creation-with-factory-method.}
   
\section{Replace Instance Variable With Local Variable}
  Review a partail view file to make sure there is no instance variable.
 
  % See the best practice details here \url{http://rails-bestpractices.com/posts/27-replace-instance-variable-with-local-variable.}
   
\section{Restrict Auto Generated Routes}
  Review a route file to make sure all auto-generated routes have corresponding actions in controller.
 
  % See the best practice details here \url{http://rails-bestpractices.com/posts/86-restrict-auto-generated-routes}
   
\section{Simplify Render In Controllers}
  Review a controller file to make sure using simplified syntax for render.
 
  % See the best practice details here \url{http://rails-bestpractices.com/posts/62-simplify-render-in-controllers.}
   
\section{Simplify Render In Views}
  Review a view file to make sure using simplified syntax for render.
 
  % See the best practice details here \url{http://rails-bestpractices.com/posts/61-simplify-render-in-views.}
   
\section{Use Before Filter}
  Review a controller file to make sure to use before\_filter to remove duplicated first code line in different action.
 
  % See the best practice detailed here \url{http://rails-bestpractices.com/posts/22-use-before_filter.}
   
\section{Use Model Association}
  review a controller file to make sure to use model association instead of foreign key id assignment.
 
  % See the best practice details here \url{http://rails-bestpractices.com/posts/2-use-model-association.}
   
\section{Use Multipart Alternative As Content Type Of Email}
  Make sure to use multipart/alternative as content\_type of email.
 
  % See the best practice details here \url{http://rails-bestpractices.com/posts/41-use-multipart-alternative-as-content\_type-of-email.}
   
\section{Use Observer}
  Make sure to use observer (sorry we only check the mailer deliver now).
 
  % See the best practice details here \url{http://rails-bestpractices.com/posts/19-use-observer.}
   
\section{Use Query Attribute}
  Make sure to use query attribute instead of nil?, blank? and present?.
 
  % See the best practice details here \url{http://rails-bestpractices.com/posts/56-use-query-attribute.}
   
\section{Use Say With Time In Migrations}
  Review a migration file to make sure to use say or say\_with\_time for customized data changes to produce a more readable output.
 
  % See the best practice details here \url{http://rails-bestpractices.com/posts/46-use-say-and-say\_with\_time-in-migrations-to-make-a-useful-migration-log.}
   
\section{Use Scope Access}
  Review a controller to make sure to use scope access instead of manually checking current\_user and redirect.
 
  % See the best practice details here \url{http://rails-bestpractices.com/posts/3-use-scope-access.}




















